Operational Feature Selection with Gaussian Mixture Models for remote sensing images

### 2016-10-07 paper.tex
Section VI.

    \begin{figure}
        \centering
        \begin{minted}[fontsize=\tiny,bgcolor=bg]{sh}
            otbcli_TrainGMMSelectionApp -io.il hyper.tif \
                            -io.vd reference.shp \
                            -gmm.varnb 20 -gmm.method forward -gmm.crit jm\
                            -gmm.best 1 -gmm.seed 0\
                            -io.out model.txt

@@

    This  second dataset  is built  from a  dataset of  remote sensing
    images distributed by the International Society for Photogrammetry
    and                         Remote                         Sensing
    (ISPRS)\footnote{\url{http://www2.isprs.org/commissions/comm3/wg4/2d-sem-label-potsdam.html}}. The
    dataset is composed of aerial images of the urban area of Potsdam.
    The area is cut into 38  patches of 6000$\times$6000 pixels with a
    resolution  of 5cm  by pixel  and 4  channels are  available: Red,
    Blue, Green  and Infrared (RGBIR).   A Digital Surface  Model with
    the same  resolution is also  provided and a  so-called normalized
    DSM representing the height above  ground. The ground-truth for 24
    tiles  are   provided  with   6  classes:  Low   vegetation,  High
    vegetation, Impervious surfaces,  Buildings, Cars, Clutter.  Three
    tiles  have  been  used  in  this  work,  they  are  displayed  in
    Figure~\ref{fig:potsdam-expl}.  Table~\ref{tab:potsdam} summarizes
    the  number of  samples of  each class.

@@

    In order to increase the dimensionality of the data, the following features are computed using the RGBIR images similarly to \cite{tuia2015multiclass}:
    \begin{itemize}
        \item 15 Radiometric indexes: NDVI, TNDVI, RVI, SAVI, TSAVI, MSAVI, MSAVI2, GEMI, IPVI, NDWI2, NDTI, RI, CI, BI, BI2~\cite{otb}.
        \item Morphological profile with reconstruction of each band with a disk of radius 5, 9, 13, 17, 21, 25, 29, 33, 37 and 41 (80 features)~\cite{fauvel2013advances};
        \item Attribute profile of each band with area as attribute and 1000, 2000, 5000, 10000 and 15000 as thresholds (40 features)~\cite{dalla2010morphological}.
        \item Attribute profile of each band with diagonal of bounding box as attribute and 100, 200, 500, 1000 and 20000 as thresholds (40 features)~\cite{dalla2010morphological}.
        \item Textural features with neighborhood of 19x19 pixels: mean, standard deviation, range and entropy (16 features)~\cite{otb}.
    \end{itemize}
    The normalized DSM and the raw  RGBIR image are added to these 191
    features and  then then  stacked to  create a  new image  with 196
    bands. %  The  training is  made with  tile 5\_11  and test  with two
    % tiles: 5\_12 which very similar with  tile 5\_11 and 3\_10 which a
    % further area a bit less urban and more industrial.

@@

    The aim  of the experiments is  to compare the proposed  method to
    some   standard  classifiers   used   in   operational  land   map
    production~\cite{rs70912356}. A non-optimized  previous version of
    the method has been already compared to other selection methods in
    \cite{fauvel2015fast}.  Hence,  the primary  objective is  to test
    the  operational  efficiency  and  it is  compared  to  other  OTB
    classifiers   used  operationally   through  their   command  line
    applications\footnote{\url{http://otbcb.readthedocs.io/en/latest/OTB-Applications.html}}.

@@

    All these classifiers are compared with 3 configurations of the proposed GMM classifier:
    \begin{itemize}
        \item one with forward selection and JM distance as criterion (GMM SFS JM);
        \item one with forward selection and Cohen's kappa as criterion (GMM SFS kappa);
        \item one with floating forward selection and JM distance as criterion (GMM SFFS JM).
    \end{itemize}

    The training set has been created  with an equal number of samples
    for each class and additionally  a spatial stratification has been
    performed, \emph{i.e.}, each training  sample belongs to a spatial
    polygon  that  does  not  intersect  spatially  with  any  spatial
    polygons used  for the validation.   Several size of  training set
    have been tested investigated.   For the Aisa dataset, experiments
    have been conducted  using 250, 500 and 1000 samples  by class and
    for the Potsdam dataset, 1000 and 50000 samples by class.

    For SFS and  SFFS selection, the number of variables to select is set to 30 for the Aisa dataset and 60 for the Potsdam dataset. Then, a small processing is made to select the number of variables used for prediction. The evolution of the criterion function is retrieved from the model file and the discrete derivative of the criterion \emph{DJ} is computed and normalized by its maximum. The subset of variables used for classification $\Omega^{*}$ is the first subset for which this value is lower than a given threshold \emph{th}, equal  to 0.001\% in this experiment,
    \begin{equation*}
        \Omega^{*} = \Omega_{\min \{k|DJ_k<th\}}
    \end{equation*}
    with $DJ_k = \frac{J(\Omega_k) - J(\Omega_{k-1})}{\max_k (J(\Omega_k) - J(\Omega_{k-1}))}$.

    The classification  rate is presented  using Cohen's kappa  but it
    has to  be noticed that the  results were  similar when looking  at the
    mean F1-score.  Processing time  has been  evaluated on  a desktop
    computer  with 4Gb  of RAM  and  Intel(R) Core(TM)  i5-3570 CPU  @
    3.40GHz $\times$ 4 processors.

@@

    The processing times of the three standard classifiers suffer from the increase of training samples. For the GMM classifier with ridge, the selection of regularization parameter is more costly with more samples because of the classification rate estimation needed. For the KNN classifier, the model stores all the training samples and the prediction implies to compute the distance to all the training samples which explains the increase of the processing time and additionally of the size of the model file. Finally, for the Random Forest classifier, the trees tends to be more deeper in order to capture the additional information available with more samples and that explains the increase of the processing time.
    % The computation of the kappa seems  to suffer from  a lack of  parallelization   which could be  improved. ->  plus tard dans la conclusion.

@@

    For the Potsdam  dataset, training samples were  selected from one
    tile (5\_11)  and validation  samples were all  the pixel  of tile
    5\_12  or  3\_10.  Table~\ref{tab:potsdam-otbsimu}  presents  the
    results.

@@

    With this second dataset, the Random Forest classifier and the GMM classifier with kappa as selection criterion distinguished themselves by good performance. When using 1000 samples, no significant difference of classification rate has been observed on test set. But, with 50,000 samples, the Random Forest classifier becomes significantly better.

@@

>>>>

    \begin{figure} 
        \centering %Pas touche l'indentation !
        \begin{minted}[fontsize=\tiny,bgcolor=bg]{sh} 
otbcli_TrainGMMSelectionApp -io.il hyper.tif \
                            -io.vd reference.shp \
                            -gmm.varnb 20 -gmm.method forward -gmm.crit jm\
                            -gmm.best 1 -gmm.seed 0\
                            -io.out model.txt

@@

    This  second dataset  is built  from a  dataset of  remote sensing
    images distributed by the International Society for Photogrammetry
    and                         Remote                         Sensing
    (ISPRS)\footnote{\url{http://www2.isprs.org/commissions/comm3/wg4/2d-sem-label-potsdam.html}}. The
    dataset is composed of aerial images of the urban area of Potsdam.
    The area  is divided  into 38  patches of  6000$\times$6000 pixels
    with a  resolution of 5cm by  pixel and 4 channels  are available:
    Red, Blue,  Green and Infrared  (RGBIR).  A Digital  Surface Model
    with  the  same  resolution  is  also  provided  and  a  so-called
    normalized  DSM   representing  the   height  above   ground.  The
    ground-truth  for  24  tiles  are provided  with  6  classes:  Low
    vegetation, High vegetation, Impervious surfaces, Buildings, Cars,
    Clutter.   Three tiles  have  been  used in  this  work, they  are
    displayed             in            Figure~\ref{fig:potsdam-expl}.
    Table~\ref{tab:potsdam} summarizes  the number of samples  of each
    class.

@@

    In order to increase the dimensionality of the data, the following features are computed using the RGBIR images similarly to \cite{tuia2015multiclass}:
    \begin{itemize}
        \item 15 Radiometric indexes: NDVI, TNDVI, RVI, SAVI, TSAVI, MSAVI, MSAVI2, GEMI, IPVI, NDWI2, NDTI, RI, CI, BI, BI2~\cite{otb}.
        \item Morphological profile with reconstruction of each band with a disk of radius 5, 9, 13, 17, 21, 25, 29, 33, 37 and 41 (80 features)~\cite{fauvel2013advances};
        \item Attribute profile of each band with area as attribute and 1000, 2000, 5000, 10000 and 15000 as thresholds (40 features)~\cite{dalla2010morphological}.
        \item Attribute profile of each band with diagonal of bounding box as attribute and 100, 200, 500, 1000 and 20000 as thresholds (40 features)~\cite{dalla2010morphological}.
        \item Textural features for  each channel with neighborhood of
          19x19 pixels:  mean, standard  deviation, range  and entropy
          (16 features)~\cite{otb}.
    \end{itemize}
    The normalized DSM and the raw  RGBIR image are added to these 191
    features and  then then  stacked to  create a  new image  with 196
    bands. %  The  training is  made with  tile 5\_11  and test  with two
    % tiles: 5\_12 which very similar with  tile 5\_11 and 3\_10 which a
    % further area a bit less urban and more industrial.

@@

    The aim  of the experiments is  to compare the proposed  method to
    some   standard  classifiers   used   in   operational  land   map
    production~\cite{rs70912356}. A non-optimized  previous version of
    the method has been already compared to other selection methods in
    \cite{fauvel2015fast}.  Hence, the primary  objective is to assess
    the  operational  efficiency  and  it is  compared  to  other  OTB
    classifiers   used  operationally   through  their   command  line
    applications\footnote{\url{http://otbcb.readthedocs.io/en/latest/OTB-Applications.html}}.

@@

    All these classifiers are compared with 3 configurations of the proposed GMM classifier:
    \begin{itemize}
    \item one with forward selection and JM distance as criterion (GMM SFS JM);
    \item one with forward selection and Cohen's kappa as criterion (GMM SFS kappa);
    \item one with floating forward selection and JM distance as criterion (GMM SFFS JM).
    \end{itemize}
    Other configurations  have been  investigated and  performs either
    equally     or     lower     in    terms     of     classification
    accuracy~\cite{al:report}. For the sake of clarity, only the three
    aforementioned configurations are discussed here.

    The training set has been created  with an equal number of samples
    for each class and additionally  a spatial stratification has been
    performed, \emph{i.e.}, each training  sample belongs to a spatial
    polygon  that  does  not  intersect  spatially  with  any  spatial
    polygons used  for the validation.   Several size of  training set
    have been  tested.  For  the Aisa  dataset, experiments  have been
    conducted using  250, 500 and  1000 samples  by class and  for the
    Potsdam dataset, 1000 and 50000 samples by class.

    For SFS and  SFFS selection, the number of variables  to select is
    set to  30 for the  Aisa dataset and  60 for the  Potsdam dataset.
    After  the selection  procedure,  the optimal  number of  selected
    variable is selected  as follow. Rather than  selecting the number
    of variables corresponding to the  highest value oft he criterion,
    the number of  retained variables is set when  the criterion stops
    to increase significantly.  It is found by  computing the discrete
    derivative of the criteria. See  %%TODO: faire une figure avec une
                                    %%courbe réelle et montrer où l'on
                                    %%s'arrete et où est le max.

    % Then, a small processing is made to select the number of variables used for prediction. The evolution of the criterion function is retrieved from the model file and the discrete derivative of the criterion \emph{DJ} is computed and normalized by its maximum. The subset of variables used for classification $\Omega^{*}$ is the first subset for which this value is lower than a given threshold \emph{th}, equal  to 0.001\% in this experiment,
    % \begin{equation*}
    %     \Omega^{*} = \Omega_{\min \{k|DJ_k<th\}}
    % \end{equation*}
    % with $DJ_k = \frac{J(\Omega_k) - J(\Omega_{k-1})}{\max_k (J(\Omega_k) - J(\Omega_{k-1}))}$.

    The    classification   rate    is    presented   using    Cohen's
    kappa. Processing  time has been  evaluated on a  desktop computer
    with  8Gb of  RAM  and  Intel(R) Core(TM)  i5-3570  CPU @  3.40GHz
    $\times$ 4 processors.

@@

    The processing times of the three standard classifiers suffer from the increase of training samples. For the GMM classifier with ridge, the selection of regularization parameter is more costly with more samples because of the classification rate estimation needed. For the KNN classifier, the model stores all the training samples and the prediction implies to compute the distance to all the training samples which explains the increase of the processing time and additionally of the size of the model file. Finally, for the Random Forest classifier, the trees tends to be deeper in order to capture the additional information available with more samples and that explains the increase of the processing time.
    % The computation of the kappa seems  to suffer from  a lack of  parallelization   which could be  improved. ->  plus tard dans la conclusion.

@@

    For the Potsdam  dataset, training samples were  selected from one
    tile (5\_11)  and validation  samples were all  the pixel  of tile
    5\_12  or  3\_10.  Table~\ref{tab:potsdam-otbsimu}  presents  the
    results in terms of classification accuracy and processing time.

@@

    With this second dataset, the Random Forest classifier and the GMM classifier with kappa as selection criterion perform the best in terms of classification accuracy. When using 1000 samples per class, no significant difference of classification rate has been observed on test set. But, with 50,000 samples per class, the Random Forest classifier becomes significantly better.

@@

working...
